\documentclass[12pt,thmsa]{article}

%\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{fullpage}
 \usepackage{color}
\usepackage{comment}
\usepackage{bm}
\usepackage{url}
\usepackage{hyperref}
\usepackage{color,hyperref}
 \definecolor{darkblue}{rgb}{0.0,0.0,0.3}
 \hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}
%%%%
\title{Problem Set 1}
\author{NAME}
\date{DATE}

\begin{document}

\doublespacing
\maketitle

\begin{enumerate}
\item Identification
\begin{enumerate}
\item Identify the following Greek letters by name. Include the meaning and common usage if you are able. 
\begin{enumerate}
\item $\delta$

\item $\beta$

\item $\alpha$

\item $\theta$

\item $\lambda$

\item $\mu$

\item $\Delta$
\end{enumerate}

\item Identify the following mathematical symbols by name. Include the meaning/usage. 
\begin{enumerate}
\item $\in$

\item $\cup$

\item $\subset$

\item $\&$ and $|$ 
\end{enumerate}
\end{enumerate}

\item Necessary and Sufficient
\begin{enumerate}
\item Provide an example of a necessary condition 

\item What is the difference between necessary and sufficient?
\end{enumerate}

\item Set theory
\begin{enumerate}
\item In roster notation, write the set characterized in set-builder notation as $S = \{x \in \mathbb{Z}, 2 <x <5\}$.

\item Is the following statement True or False? $\forall x \in S, x\geq 2$ for $S = \{3,2,5,9\}$.

\item Is the following statement True or False? $\exists x \in S\; s.t. \; x \notin \mathbb{Z}$ for $S = \{3,2,5,9\}$.

\item Is $\{1,2,3,4\}$ a subset of $\{4,3,1,2\}$? Is it a proper subset?

\item Using logical symbols (including $\exists$ and $\forall$) write the definition of a proper subset.

\item If $A = \{soup, 8\}$ and $B = \{x, soup\}$ find $A \cup B$.

\item (Follow-up): Now find $A \cap B$.

\item (Follow-up): Find the Cartesian Product $A \times B$.
\end{enumerate}

\item Topics and Questions
    \begin{enumerate}
     \item List three things you struggled with on today's assignment. 

     \item What is your plan for improving the items listed above?

     \item What percent of the material was new to you today?

     \item What is one new concept you learned today?

     \item What question do you still have about the material?
    \end{enumerate}

\end{enumerate}
\end{document}