\documentclass[12pt,thmsa]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{fullpage}
 \usepackage{color}
\usepackage{comment}
\usepackage{bm}
\usepackage{url}
\usepackage{hyperref}
\usepackage{color,hyperref}
 \definecolor{darkblue}{rgb}{0.0,0.0,0.3}
 \hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}
%%%%
\title{Problem Set 3}
\author{NAME}
\date{DATE}

\begin{document}

\maketitle 

Due on Friday!   

\begin{enumerate}

    \item Summation
        \begin{enumerate}
            \item $\sum_{n=1}^{7}3$
            \item $\sum_{n=0}^{4} 2n+8$
        \end{enumerate}

    \item Limits and Continuity
        \begin{enumerate}
            \item What is a limit?
            \item Why do we care about limits and tangents?
            \item What is a continuous function?
        \end{enumerate}

\item Derivative Foundations
\begin{enumerate}
\item What is a derivative? Why might we find it useful? 
\item Find an example on the internet of a derivative used to solve a problem in real life. 
\end{enumerate}

\item Derivatives by hand : Compute manually, using the formula for a derivative (i.e. $ lim_{h\rightarrow0}\frac{f(x+h)-f(x)}{h}$). 
\begin{enumerate}
\item  $f(x) =x^3$
\end{enumerate}

\item Derivatives using the formulas: Compute these using the rules from class. 
\begin{enumerate}
\item  $f(x) = 2x^2 + 7x + 9$ 
\item  $f(x) = 3x^2$
\item  $f(x) =x^3$
\item  $f(x)=ln(x)$
\item $f(x)=e^{3x}$
\item $f(x)=2e^{-2x}-x^{0.5}$
\item $f(x)=(3x^4-6x+2)(x^2-4)$
\item $f(x)=ln(e^x)$
\end{enumerate}

\item Check out this resource on critical points: 
\url{https://www.cuemath.com/calculus/critical-point/}. Based on this, what are critical points and how do derivatives help us find them? Why might this be important (can provide an example from another website)? 



\item Topics and Questions
    \begin{enumerate}
        \item List three things you struggled with on today's assignment. 
        \item What is your plan for improving the items listed above?
        \item What percent of the material was new to you today?
        \item What is one new concept you learned today?
        \item What question do you still have about the material?
    \end{enumerate}


\end{enumerate}


\end{document}

